On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification

### 2020-08-13 paper/cav_determinisim_main.tex
checked link between deterministic threshold and tolerance

\begin{abstract}
\noindent 
Game engines are increasingly used as simulation platforms by the autonomous vehicle (AV) community to develop vehicle control systems and test environments. 
%
Determinism is a key requirement for simulation-based development and verification; it ensures tests are repeatable, so that software bugs can be found and fixed, and simulation results are trustworthy.
%
% KIE: Ideally, these engines are deterministic, but we've found that they are not. :(
% These engines must be deterministic for finding and resolving defects (bugs) in software and ensuring auditable and repeatable tests can be executed for verification and validation (V\&V). 
%
However, game engines do not need to be deterministic per se. % when used for games. 
%
This paper identifies the requirements these engines need to fulfil when used for AV development and testing, and proposes a method to determine the \textit{simulation variance} for specific configurations and utilisation of the software and hardware, respectively. 
%
A case study using CARLA, an open-source simulation environment powered by Unreal Engine, illustrates the proposed method and results from an example scenario indicate the presence of a deterministic operational simulation setup. \todo{Review again at end. Potentially give away more on results to motivate the reader to continue.}

@@

@@

The autonomous vehicle community are adopting game engine simulators for such purposes as control system development and verification. Having a deterministic simulator is required for reliable and repeatable results, to find and fix software bugs and ensure results are trustworthy. If a simulator is non-deterministic then practitioners should at least be aware of, and how to find, the operational domains for reliable results. During an investigation into the CARLA simulator, a \textit{simulation variance} was observed for repeated tests with the same initial conditions indicating non-deterministic execution. During the study we hypothesized and uncovered several parameters that contribute towards greater simulation variance and hence non-deterministic execution, such as actor collisions and system resource utilisation. The results of the investigation are hardware and software version specific, so we present a universal methodology that can be used to find the \textit{deterministic operational domain} for any given system configuration. This methodology will allow the AV verification community using simulation tools to ensure their results are reliable and trustworthy.

>>>>

\begin{abstract}
\noindent 
Game engines are increasingly used as simulation platforms by the autonomous vehicle (AV) community to develop vehicle control systems and test environments. 
%
Determinism is a key requirement for simulation-based development and verification; it ensures tests are repeatable, so that software bugs can be found and fixed, and simulation results are trustworthy.
\todo[inline]{Determinism is a key requirement for simulation-based development and verification; it ensures the simulation is precise and that tests are repeatable, so that software bugs can be found and fixed, and simulation results are trustworthy.}
%
% KIE: Ideally, these engines are deterministic, but we've found that they are not. :(
% These engines must be deterministic for finding and resolving defects (bugs) in software and ensuring auditable and repeatable tests can be executed for verification and validation (V\&V). 
%
However, game engines do not need to be deterministic per se. % when used for games. 
%
This paper identifies the requirements these engines need to fulfil when used for AV development and testing, and proposes a method to determine the \textit{simulation variance} for specific configurations and utilisation of the software and hardware, respectively. 
%
A case study using CARLA, an open-source simulation environment powered by Unreal Engine, illustrates the proposed method and results from an example scenario indicate the presence of a deterministic operational simulation setup. \todo{Review again at end. Potentially give away more on results to motivate the reader to continue.}

@@

\todo[inline]{To ensure precise simulation outcomes the physics setting $dt$ must be fixed, along with any actor navigation meshes, random number seeds, game engine setup and simulation specific parameters.}

@@

The autonomous vehicle community are adopting game engine simulators for such purposes as control system development and verification. Having a deterministic simulator is required for reliable and repeatable results, \todo[inline]{Having a deterministic simulator is required for precise results,} to find and fix software bugs and ensure results are trustworthy. If a simulator is non-deterministic then practitioners should at least be aware of, and how to find, the operational domains for reliable results. During an investigation into the CARLA simulator, a \textit{simulation variance} was observed for repeated tests with the same initial conditions indicating non-deterministic execution. During the study we hypothesized and uncovered several parameters that contribute towards greater simulation variance and hence non-deterministic execution, such as actor collisions and system resource utilisation. The results of the investigation are hardware and software version specific, so we present a universal methodology that can be used to find the \textit{deterministic operational domain} for any given system configuration. This methodology will allow the AV verification community using simulation tools to ensure their results are reliable and trustworthy.
